Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-11-22 paper.tex
First cut at an intro

\title{Performance Evaluation of Big Data Management Strategies for Neuroinformatics}

@@

%
% Abstract
%
\begin{abstract}

@@

%
% Section
%
\section{Introduction} % 1 page with abstract

% Big Data Engines have been developed mostly for text applications, 
%  in particular Web. Mention Spark and Dask
Big Data processing engines have been leveraging data locality, 
in-memory computing and lazy evaluation to reduce the impact of data 
manipulation on the performance of applications. From early Hadoop 
implementations~\cite{hadoop} to Apache Spark~\cite{spark} and 
Dask~\cite{dask}, these frameworks have gained popularity and are now 
becoming mainstream in data analytics.

In scientific data analysis, however, the adoption of Big Data engines

% Scientific applications, including neuroinformatics are different: CLI, binary data, shared HPC clusters (although ref to cloud).
% In neuroinformatics, nipype, psom, etc
% This follows the current trend to big data science, data sets *and derived data* are larger (give examples).

% Our question: is it relevant to use Big Data engines for neuroinformatics? 
% Implications: migrate existing pipelines, adopt a new framework (costly)
% Some papers said yes, for instance Ariel's, Thunder and the Spark provenance one (https://github.com/UFFeScience/SAMbA).
% Perhaps this too: https://www.sciencedirect.com/science/article/pii/S0925231218311470 Spark is being used to implement neuroinformatics pipelines, but is it
% the right tool?
% But a quantitative evaluation is lacking.

% More precisely, we intend to quantify the effect of in-memory computing, data locality, lazy evaluation
% on various use cases. We intentionally leave scheduling out here.

% The use of a Big Data engine is not obvious when tmpfs and asynchronous writes
% are available and could be used to emulate in-memory computing. Our goal is to quantify when and how Big Data engines will be useful.
% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).

% Talk about pilot jobs and say we don't consider them here.

% define pipeline and workflows

% Here we focus on performance but we are aware of other desirable properties of workflow engines, including provenacne.

% Outline

% Talk about containers somewhere.

\section{Materials and Methods} % 4 pages

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless, which we will use to define our benchmarks.

% Cache eviction: LRU/n.

@@

@@

@@

% Talk about ~\cite{lebre2015adding}. Most simulation toolkits have focused on task 
%  scheduling. And https://dl.acm.org/citation.cfm?id=3041715

% And Wrench.

@@

@@

@@

@@

\section{Results} % 2 pages

@@

\subsection{CPU Time}
% Comment on the general trend

@@

\section{Discussion} % 2 pages

\subsection{Effect of In-Memory Computing}
% Shared fs vs local disk vs in-memory: what do we gain?

\subsection{Effect of Data Locality}

@@

\section{Conclusion} % 1 page with refs

% When is it useful to use a Big Data engine?

@@

>>>>

\title{Performance Evaluation of Big Data Processing Strategies for Neuroimaging}

@@

\begin{abstract}

@@

% Talk about containers somewhere?

\section{Introduction} % 1 page with abstract

Big Data processing engines leverage data locality, 
in-memory computing, and lazy evaluation to reduce the impact of data 
manipulation on the performance of applications. From early MapReduce 
implementations~\cite{dean2008mapreduce} to Apache Spark~\cite{zaharia2016apache} and now 
Dask~\cite{rocklin2015dask}, these frameworks have become 
mainstream tools for data analytics. Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
However, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus}, 
Taverna~\cite{oinn2004taverna}, and many others.

Scientific applications slightly differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (Web search, frequent pattern mining, recommender systems, 
etc~\cite{leskovec2014mining}), implemented in consistent software libraries, and 
running on clouds or dedicated commodity clusters with locality-aware 
file systems such as HDFS~\cite{shvachko2010hadoop}, scientific applications often 
involve binary data such as images and signals, processed by 
command-line tools using a mix of programming languages (C, Fortran, 
Python, shell scripts) and deployed on large, shared clusters where 
data is fetched from data nodes to compute nodes using shared file 
systems such as Lustre~\cite{schwan2003lustre}. Such differences in applications 
and infrastructures have important consequences. To mention only one, 
in-memory computing requires instrumentation to be applied to 
command-line tools.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(\texttt{tmpfs}) and memory-mapped files\todo{check that} might also 
explain the lack of adoption of Big Data engines for scientific 
applications.

In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

% https://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20130001600.pdf

Neuroimaging, our primary field of interest, is no exception to the 
generalized rise of data volumes in science, due to the joint increase 
of image resolution and subject cohort sizes~\cite{van2014human}. 
Processing engines have been developed with neuroinformatics 
applications in mind, for instance Nipype~\cite{gorgolewski2011nipype} 
or PSOM~\cite{bellec2012pipeline}. Big Data engines have been used for 
neuroimaging applications too, including in the Thunder 
project~\cite{freeman2014mapping} and in more specific works such 
as~\cite{makkie2019fast}. However, no quantitative performance 
evaluation has been conducted on neuroimaging applications to assess the 
added-value of Big Data engines compared to traditional processing engines.

This paper addresses the following questions:
\begin{enumerate}
\item What is the effect of in-memory computing and data locality on neuroimaging applications?
\item Can in-memory computing be effectively enabled by the operating system rather than the data processing engine?
\end{enumerate}

Answers to these questions have important implications. 
In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
presented. It  concludes that these systems need to be extended to 
better address scientific code integration, data partitioning, data 
formats and system tuning. The work in~\cite{samba} also extended Spark 
with a better provenance system for scientific applications. We argue 
that such efforts should only be conducted if substantial performance 
improvements are expected from in-memory computing, lazy 
evaluation\todo{are we talking about lazy evaluation at all?} or data 
locality. On the other hand, scientific data processing engines are 
still being developed, and it is legitimate to wonder whether these 
projects should just be migrated to Spark, Dask, or other Big Data 
engines.

Our study focuses on performance. We intentionally do not intend to 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, provenance capture and 
representation, portability or reproducibility, which are otherwise 
critical concerns. Besides, our study of performance focuses on the 
impact of data writes and transfers. It purposely leaves out task 
scheduling to computing resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion though.

\todo{Perhaps talk about modeling and simulation here.}

Our methods, including performance models, processing engines, \todo{Make this paragraph more informative}
applications and infrastructure used are described in 
Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
results which we discuss in Section~\ref{sec:discussion}. 
Section~\ref{sec:conclusion} concludes on the answer to our two research questions.

\section{Materials and Methods} % 4 pages
\label{sec:methods}

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless, which we will use to define our benchmarks. It 
should be noted that leveraging the page cache, and therefore ensuring 
that Equation~\ref{eq:page-cache-inequality} holds, has important 
performance implications: with page caching, the application will 
experience an almost negligible write time, while without page caching 
the write throughput will be that of the disk.

% Cache eviction: LRU/n.
% page caching is mentioned in https://ntrs.nasa.gov/archive/nasa/casi.ntrs.nasa.gov/20130001600.pdf too
% Note: writeback is dangerous as a system failure could result in data loss.

@@

%~ In any case, I think we want the following inequality to hold:
%~ ```
%~ D/C <= \Delta / \Gamma
%~ ```
%~ where Delta is now the apparent bandwidth of the file system and Gamma 
%~ is the max number of concurrent tasks in the cluster. Whether page 
%~ caching is used or not will affect what happens when the inequality 
%~ holds: if page caching is here, write time will be 0, otherwise it will 
%~ be non-zero. If the inequality doesn't hold then we have congestion and 
%~ it's a different game: transfer times might tend to infinity and data 
%~ loss is to be expected.

% Say that we don't use HDFS because we focus on clusters and HDFS on multi-tenant clusters is difficult (refer to hadoop on demand).

@@

In~\cite{saini2012performance} the Lustre file system is benchmarked on NASA applications.

@@

% Talk about ~\cite{lebre2015adding} and \cite{wrench}. Most simulation toolkits have focused on task 
%  scheduling. And https://dl.acm.org/citation.cfm?id=3041715

\subsection{Task Scheduling}

% Explain how we ensured that both engines lead to the same scheduling.
% Talk about pilot jobs and say we don't consider them here.

@@

\begin{table}
\centering
\begin{tabular}{c|c}
$\delta/\gamma$ & $\Delta/\Gamma$\\
\hline
50~MB/s         & 5~MB/s
\end{tabular}
\label{table:infrastructure}
\caption{Bandwidth thresholds on the target cluster}
\end{table}

@@

% Talk about our I/O pattern: random I/O, sequential.

\begin{table}
\centering
\begin{tabular}{c|c}
Application & $\frac{D}{C}$ (GB/s)\\
\hline
Big Brain            & [0-75] \\
fMRI pre-processing  & [0-1.5]
\end{tabular}
\label{table:applications}
\caption{Data-compute ratios for typical neuroinformatics applications}
\end{table}
% fMRI: about 100M in, 2GB out,2 30 min processing time to 10 hours

@@

We tuned job scheduling to balance the load among
cluster nodes. This is the default behavior of Spark, but required
specific instrumentation in Nipype.

@@

\section{Results} % 2 pages
\label{sec:results}

@@

\subsection{CPU Time}
% Comment on the general trend
% We can't really say that imaging tasks are not small. Even Freesurfer could be decomposed.
%

@@

\section{Discussion} % 2 pages
\label{sec:discussion}
\subsection{Effect of In-Memory Computing}
% Shared fs vs local disk vs in-memory: what do we gain?

\subsection{Effect of Data Locality}
% It should be important, in particular when there is contention and when
% tasks are small.

@@

\section{Conclusion} % 1 page with refs
\label{sec:conclusion}
% When is it useful to use a Big Data engine?

@@

\todo{Acknowledge Dell and ask them if they want to co-author the paper.}
